---
title: "String Manipulation with string R"
output: html_notebook
---

## Quote

First, you should prefer double quotes (") to single quotes ('). That means, whenever you are defining a string your first intuition should be to use ".

Unfortunately if your string has " inside it, R will interpret the double quote as "this is the end of the string", not as "this is the character "". This is one time you can forget the first guideline and use the single quote, ', to define the string.

You can escape quotes inside strings using a backslash (\").


```{r, eval= FALSE}

# Define line1
line1 <- "The table was a large one, but the three were all crowded together at one corner of it:"

# Define line2
line2 <- "\"No room! No room!\" they cried out when they saw Alice coming."

# Define line3
line3 <- "\"There's plenty of room!\" said Alice indignantly, and she sat down in a large arm-chair at one end of the table."

```

## writeLines()


```{r, eval = FALSE}

# Putting lines in a vector
lines <- c(line1, line2, line3)

# Print lines
print(lines)

# Use writeLines() on lines
writeLines(lines)

# Write lines with a space separator
writeLines(lines, sep = " ")


# Use writeLines() on the string "hello\n\U1F30D"
writeLines("hello\n\U1F30D")

```

## Escape sequences

A sequence in a string that starts with a ``\`` is called an escape sequence and allows us to include special characters in our strings. You saw one escape sequence in the first exercise: ``\"`` is used to denote a double quote.

In ``"hello\n\U1F30D"`` there are two escape sequences: ``\n`` gives a newline, and ``\U`` followed by up to 8 hex digits sequence denotes a particular Unicode character.

Unicode is a standard for representing characters that might not be on your keyboard. Each available character has a Unicode code point: a number that uniquely identifies it. These code points are generally written in hex notation, that is, using base 16 and the digits 0-9 and A-F. You can find the code point for a particular character by looking up a code chart. If you only need four digits for the codepoint, an alternative escape sequence is ``\u``.

When R comes across a ``\`` it assumes you are starting an escape, so if you actually need a backslash in your string you'll need the sequence ``\\``.


```{r, eval = FALSE}

# Should display: To have a \ you need \\
writeLines("To have a \\ you need \\\\")

# Should display: 
# This is a really 
# really really 
# long string
writeLines("This is a really really really long string")

# Use writeLines() with 
# "\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e"

writeLines("\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e")

```


## Turning numbers into strings
### format() and formatC()

```{r, eval = FALSE}

# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1), digits = 1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1), digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digits = 2)

# Format income to whole numbers
format(income, digits = 2)

# Format p_values in fixed format

format(p_values, scientific = FALSE)


```

## Controlling other aspects of the string with format() and formatC()

```{r, eval = FALSE}

formatted_income <- format(income, digits = 2)

# Print formatted_income
print(formatted_income)
# Call writeLines() on the formatted income
writeLines(formatted_income)
# Define trimmed_income
trimmed_income<- format(income, trim = TRUE, digits = 2)

# Call writeLines() on the trimmed_income
writeLines(trimmed_income)

# Define pretty_income
pretty_income<- format(income, digits = 2, big.mark = ",")

# Call writeLines() on the pretty_income
writeLines(pretty_income)

######## formatC()#############

# From the format() exercise
x <- c(0.0011, 0.011, 1)
y <- c(1.0011, 2.011, 1)

# formatC() on x with format = "f", digits = 1
formatC(x, format = "f", digits = 1)

# formatC() on y with format = "f", digits = 1
formatC(y, format = "f", digits = 1)

# Format percent_change to one place after the decimal point
formatC(percent_change, format = "f", digits = 1)

# percent_change with flag = "+"
formatC(percent_change, format = "f", flag = "+", digits = 1)

# Format p_values using format = "g" and digits = 2

formatC(p_values, format = "g", digits = 2)

```

## Putting strings together

### paste()


```{r, eval = FALSE}

# Add $ to pretty_income
paste("$", pretty_income, sep = "")
 
# Add % to pretty_percent
paste(pretty_percent, "%", sep = "")


# Create vector with elements like 2010: +4.0%`
year_percent <- paste(years,": ",pretty_percent,"%", sep ="")

# Collapse all years into single string
paste(year_percent, collapse = ", ")

```


## Creating a table

```{r, eval = FALSE}

# Define the names vector
income_names <- c("Year 0", "Year 1", "Year 2", "Project Lifetime")

# Create pretty_income
pretty_income <- format(income, digits = 2, big.mark = ",")

# Create dollar_income
dollar_income <- paste("$", pretty_income, sep = "")

# Create formatted_names
formatted_names <- format(income_names, justify = "right")

# Create rows
rows<- paste(formatted_names, dollar_income, sep="   ")

# Write rows
writeLines(rows)


```


## Let's order pizza!


```{r, eval = FALSE}


# Randomly sample 3 toppings
my_toppings <- sample(toppings, size = 3)

# Print my_toppings
my_toppings

# Paste "and " to last element: my_toppings_and
my_toppings_and <- paste(c("", "", "and "), my_toppings, sep = "")

# Collapse with comma space: these_toppings
these_toppings <- paste(my_toppings_and, collapse = ", ")

# Add rest of sentence: my_order
my_order <- paste("I want to order a pizza with ", these_toppings, ".", sep = "")

# Order pizza with writeLines()
writeLines(my_order)


```



















































































































































































































































