---
title: "Visible Aesthetics"
output: html_notebook
---

## Aesthetics

![Aesthetics](D:/Documents/R_Projects/Data_Camp_Tutorials/ggplot()/Aesthetics.JPG)


```{r}
# 1 - Map mpg to x and cyl to y
ggplot(mtcars, aes(x = mpg, y = cyl)) +
  geom_point()
  
# 2 - Reverse: Map cyl to x and mpg to y
ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()

# 3 - Map wt to x, mpg to y and cyl to col
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point()

# 4 - Change shape and size of the points in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(shape = 1, size =4)

```

## All about aesthetics

```{r}

# am and cyl are factors, wt is numeric
class(mtcars$am)
class(mtcars$cyl)
class(mtcars$wt)

# From the previous exercise
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(shape = 1, size = 4)

# 1 - Map cyl to fill

ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
geom_point(shape = 1, size = 4)
  
# 2 - Change shape and alpha of the points in the above plot

ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)

# 3 - Map am to col in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl, col = am)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)



```


## All about aesthetics part 2

```{r}

# Map cyl to size

ggplot(mtcars, aes(x = wt, y = mpg,size = cyl))+geom_point()

# Map cyl to alpha

ggplot(mtcars, aes(x = wt, y = mpg,alpha = cyl))+geom_point()
# Map cyl to shape 

ggplot(mtcars, aes(x = wt, y = mpg,shape = cyl))+geom_point()



# Map cyl to label

ggplot(mtcars, aes(x = wt, y = mpg, label = cyl))+geom_point()+geom_text()


```


## All about attributes, part 1

```{r}
# Define a hexadecimal color
my_color <- "#4ABEFF"

# 1 - First scatter plot, with col aesthetic:

ggplot(mtcars, aes(x =wt, y = mpg, col = cyl))+geom_point(col = my_color)


# 2 - Plot 1, but set col attributes in geom layer:

ggplot(mtcars, aes(x =wt, y = mpg, col = cyl))+geom_point(col = my_color)


# 3 - Plot 2, with fill instead of col aesthetic, plut shape and size attributes in geom layer.
ggplot(mtcars, aes(x =wt, y = mpg, fill = cyl))+geom_point(col = my_color, size = 10, shape = 23)

```

## All about attributes, part 2


```{r}
# Expand to draw points with alpha 0.5
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl))+geom_point(alpha = 0.5) 

  
# Expand to draw points with shape 24 and color yellow
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +geom_point(shape = 24, col = "yellow")

  
# Expand to draw text with label rownames(mtcars) and color red
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl))+geom_text(aes(label = rownames(mtcars)), col ="red")



```

## Going all out


```{r}
# Map mpg onto x, qsec onto y and factor(cyl) onto col
ggplot(mtcars, aes(x= mpg, y = qsec, col = factor(cyl)))+geom_point()


# Add mapping: factor(am) onto shape
ggplot(mtcars, aes(x= mpg, y = qsec, col = factor(cyl),shape=factor(am)))+geom_point()

# Add mapping: (hp/wt) onto size

ggplot(mtcars, aes(x= mpg, y = qsec, col = factor(cyl),shape=factor(am),size = (hp/wt)))+geom_point()

```

## Positions

![Positions](D:/Documents/R_Projects/Data_Camp_Tutorials/ggplot()/position.JPG)


```{r, eval=FALSE}
# The base layer, cyl.am, is available for you
# Add geom (position = "stack" by default)
cyl.am<- ggplot(mtcars, aes(x = cyl, y = factor(am))
cyl.am + 
  geom_bar()

# Fill - show proportion
cyl.am + 
  geom_bar(position = "fill")  

# Dodging - principles of similarity and proximity
cyl.am +
  geom_bar(position = "dodge") 

# Clean up the axes with scale_ functions
val = c("#E41A1C", "#377EB8")
lab = c("Manual", "Automatic")
cyl.am +
  geom_bar(position = "dodge") +
  scale_x_discrete("Cylinders") + 
  scale_y_continuous("Number") +
  scale_fill_manual("Transmission", 
                    values = val,
                    labels = lab) 


```


## Setting a dummy aesthetic


```{r}

# 1 - Create jittered plot of mtcars, mpg onto x, 0 onto y
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter()

# 2 - Add function to change y axis limits

ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter()+scale_y_continuous(limits = c(-2,2))
```

## Overplotting 1 - Point shape and transparency

```{r}
# Basic scatter plot: wt on x-axis and mpg on y-axis; map cyl to col
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl))+geom_point(size = 4)


# Hollow circles - an improvement
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl))+geom_point(size = 4, shape = 1)



# Add transparency - very nice

ggplot(mtcars, aes(x = wt, y = mpg, col = cyl))+geom_point(size = 4,alpha = 0.6)

```


## Overplotting 2 - alpha with large datasets

```{r}

# Scatter plot: carat (x), price (y), clarity (color)

ggplot(diamonds, aes(x = carat, y = price, color= clarity))+ geom_point()

# Adjust for overplotting

ggplot(diamonds, aes(x = carat, y = price, color = clarity))+ geom_point(alpha = 0.5)


# Scatter plot: clarity (x), carat (y), price (color)

ggplot(diamonds, aes(x = clarity, y = carat, color = price))+ geom_point(alpha = 0.5)


# Dot plot with jittering

ggplot(diamonds, aes(x = clarity, y = carat, color = price))+ geom_point(alpha = 0.5, position = "jitter")

```






















